---
title: "Regression Analysis with Shiny App"
author: "tuhin mahmud"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    fig_caption: yes
    keep_md: yes
runtime: shiny
---
#Synopsis
The purpose of this project is to look at a data set of a collection of cars.We are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). We want to answer following two questions:

1. "Is an automatic or manual transmission better for MPG"
1. "Quantify the MPG difference between automatic and manual transmissions"

##Summary
Studying ther results of the linear regression and multi variable linear regression we find that automatic transmission is better for MPG but impact of transmission type on MPG is lowered when considered with other variables.

For linear regression the MPG increasses  7.245 MPG , whereas if we consider other important variables like cylinder type , weight  and horse power the increase in MPG is about 1.809 MPG.


```{r echo=FALSE}
##clean data
library(datasets)
data<-mtcars
data$cyl<-factor(data$cyl)
data$am<-factor(data$am)
levels(data$am)<-c("automatic","manual")
data$gear<-factor(data$gear)
data$carb<-factor(data$carb)
```

## Build and  Analyze Linear Regression Models
###*Shiny app* implementtion
####*Input* : Choose a regression Model from the drop down selection.
####*Ouput*  :Scatter plot and regression line for analysis

```{r results="hide"}
model1<-lm(mpg~am,data=data)
model2<-lm(mpg~cyl,data=data)
model3<-lm(mpg~cyl+wt,data=data)
model4<-lm(mpg~cyl+wt+hp,data=data)
model5<-lm(mpg~am + cyl + wt + hp,data=data)
model6<-lm(mpg~.,data=data)
model7<-lm(mpg~ wt + qsec + am,data=data)

```

```{r echo=FALSE}
library(ggplot2)
library(gridExtra)
```

```{r echo=FALSE}
library(ggplot2)
shinyApp(
ui=fluidPage(
  titlePanel("Regression Model"),
  sidebarLayout(
    sidebarPanel(
      selectInput("model.name", "Choose a model:", choices = c("model1", "model2","model3","model4","model5","model6","model7"))
    ),
    mainPanel(   
      plotOutput("regressionPlot")     
    )
  )
),
server=function(input, output) { 
  output$regressionPlot <- renderPlot({
     
  par(cex=.8)
  attach(data)
  if( input$model.name=="model1") {
      plot(am, mpg, main="Scatterplot mpg vs am and regression line lm<-mpg~am", 
         xlab="Automatic or Manula ", ylab="Miles Per Gallon ", pch=19)
      abline(model1)
      p <- ggplot(mtcars, aes(wt, mpg)) + 
           geom_point(aes(colour = factor(cyl)), size = 4)+
           geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x)
  } else if ( input$model.name=="model2") {
     p <- ggplot(data=mtcars, aes( x=cyl, y=mpg)) + 
          geom_point(aes(colour = factor(cyl)), size = 4)+
          geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x)
      #plot(wt, mpg, main="Scatterplot mpg vs wt and and regression line lm<-mpg~cyl", 
      #xlab="Car weight ", ylab="Miles Per Gallon ", pch=19)
      #abline(model2)
  } else if ( input$model.name=="model3") {
      p <- ggplot(data=mtcars, aes( x=cyl+wt, y=mpg)) + 
           geom_point(aes(colour = factor(cyl)), size = 4)+
           geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x)
      #plot(wt, mpg,main="Scatterplot mpg vs wt and and regression line lm<-mpg~cyl+wt", 
      #xlab="Car weight ", ylab="Miles Per Gallon ", pch=19)
      #abline(model3)
  }else if ( input$model.name=="model4") {
      p <- ggplot(data=mtcars, aes( x=cyl+wt+hp, y=mpg)) +
          geom_point(aes(colour = factor(cyl)), size = 4)+
          geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x)
      #plot(wt, mpg,main="Scatterplot mpg vs wt  and regression line lm<-mpg~cyl", 
      #xlab="Car weight ", ylab="Miles Per Gallon ", pch=19)
      #abline(model4)
  } else if ( input$model.name=="model5") {
      
      p <- ggplot(data=mtcars, aes( x=am+cyl+wt+hp, y=mpg)) + 
          geom_point(aes(colour = factor(am)), size = 4)+
          geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x)
      #plot(wt, mpg,main="Scatterplot mpg vs wt and regression line lm<-am + cyl + wt + hp", 
      #xlab="Car weight ", ylab="Miles Per Gallon ", pch=19)
      #abline(model5)
  } else  if ( input$model.name=="model6") {
      
      p <- ggplot(data=mtcars, aes( x=cyl+disp+hp+drat+wt+qsec+vs+am+gear+carb, y=mpg)) + 
          geom_point(aes(colour = factor(am)), size = 4)+
          geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x)
      #plot(wt, mpg,main="Scatterplot mpg vs wt  and regression line lm<-mpg~.", 
      #xlab="Car weight ", ylab="Miles Per Gallon ", pch=19)
      #abline(model6)
  }else  if ( input$model.name=="model7") {      
      p <- ggplot(data=mtcars, aes( x=wt + qsec + am, y=mpg)) + 
          geom_point(aes(colour = factor(am)), size = 4)+
          geom_smooth(method = "lm", se=FALSE, color="black", formula = y ~ x)
      #plot(wt, mpg,main="Scatterplot mpg vs wt and regression line lm<-mpg~wt + qsec + am", 
      #xlab="Car weight ", ylab="Miles Per Gallon ", pch=19)
      #abline(model7)
  }
  print(p)
   })
 }
)
```
  
```{r}
#Pair wise anova comparison of the models
anova(model1,model2)
anova(model1,model3)
anova(model1,model4)
anova(model1,model5)
anova(model1,model6)
anova(model1,model7)
# http://people.duke.edu/~rnau/Notes_on_linear_regression_analysis--Robert_Nau.pdf
# Comparing regression model using R squared metric
aR2<-c(summary(model1)$adj.r.squared,
          summary(model2)$adj.r.squared,
          summary(model3)$adj.r.squared,
          summary(model4)$adj.r.squared,
          summary(model5)$adj.r.squared,
          summary(model6)$adj.r.squared,
          summary(model7)$adj.r.squared
          )
names(aR2)<-c("model1","model2","model3","model4","model5","model6","model7")
print(aR2)  
 
```
#Variable Selection 
using method suggested in http://www.statmethods.net/stats/regression.html
```{r}
# Stepwise Regression
library(MASS)
fit <- lm(mpg~am+cyl+wt+hp,data=data)
step <- stepAIC(fit, direction="both")
step$anova # display results


```
# Multvariable Regression Analysis
```{r results="hide"}
sm = step(lm(data = mtcars, mpg ~ .))
```
```{r}
summary(sm)
```
#Results

From the Adjusted R<sup>2</sup> results and study of the anova and also as suggested by Stepwise Regression study Model5 turns out to be the best fit.

Model has Adjusted R-squared value of  0.84 with p-value 1.51e-10, so we will need to reject the null hypothesis and propse model5 is better.
```{r}
model5$coefficients[2]
```
From the above best fit model and the Esitmate coefficient of "am manual" is 
1.8092, which is the quantity of impact of transmission type on MPG.
```{r}
model1$coefficients[2]
```
Similar coefficient of single varialbe linear regression model1 shows the impact of the transmission type is 7.245.

```{r }
#summary(model5)
#str(model5)
layout(matrix(c(1,2,3,4),2,2)) 
plot(model5)

```
